---
title: Generative Art with Jasmines
author: Randi Bolt
date: '2021-10-19'
slug: generative-art-with-jasmines
categories:
  - rstudio
  - art
tags:
  - rstudio
  - tutorial
  - art
  - jasmines
  - dplyr
  - design
---

For a class fall term 2021 at PSU I had to do a piece of 'self' art, so to expland my skills in R I decided to try out some generative art. Most of these are based off example given on djnavarro's github [here](https://github.com/djnavarro/jasmines). 

### Final Design 

```{r, message=F, error=F}
library(dplyr) 
library(jasmines)
use_seed(5) %>%
  entity_circle(grain = 1000, size = 10) %>%
  unfold_warp(iterations = 100) %>%
  style_ribbon(
    color = "#9E4244",
    background = "#FDAB9F")
```

When playing around with this package, I initially had something less fluid and full of right angles, but wanted to show more movement in the design. I have a dance background and aside from fluid movement we also focused a lot on circles and rotation. The other reason I like this design is because it reminds me of a flower. I have seven tattoos, two of which are flowers, and I’d like to add more flowers someday. The two colors I choose are salmon and rosewood. I really love different shades of pink, and colors like salmon, and rosewood feel like a more sophisticated pink to me. 

### Other Designs
```{r}
use_seed(1) %>%
  entity_circle(grain = 1000, size = 4) %>%
  unfold_warp(iterations = 100) %>%
  style_ribbon(
    palette="base",
    colour = "ind",
    background = "mistyrose")
```

### Typophobia

```{r}
library(jasmines)
  scene_discs(
    rings = 13, 
    points = 500, 
    size = 5
  ) %>%
  mutate(ind = 1:n()) %>%
  unfold_warp(
    iterations = 10,
    scale = .5, 
    output = "layer" 
  ) %>%
  unfold_tempest(
    iterations = 10,
    scale = .01
  ) %>%
  style_ribbon(
    color = "#48AAAD",
    colour = "ind",
    alpha = c(.1,.1),
    background = "#016064" 
  ) 
```

### Snake Charmer

```{r}
use_seed(4) %>%
  entity_circle(grain = 10000) %>%
  unfold_tempest(iterations = 13) %>%
  style_ribbon(background = "oldlace")
```


