---
title: Transformations and Weighting to Correct Model Inadequacies
author: Randi Bolt
date: '2021-11-29'
slug: transformations-and-weighting-to-correct-model-inadequacies
categories:
  - statistics
tags:
  - book problems
  - data
  - notes
  - multiple linear regression
  - statistics
  - tutorial
  - transformations
  - weighting to correct model inadeqacies
---

```{r, message=FALSE, echo=F}
# libraries
library(latex2exp) # letex on graphs 
library(magrittr) # %>%
library(dplyr) # arrange
library(MASS) # boxcox()
```

# Introduction 

The examples are from [this] textbook, and my class notes are [here]. 

# Example 5.1 : The Electric Utility Data

An electric utility company is interested in developing a model relating peak-hour demand $(y)$ to total energy usage during the month $(x)$. 

```{r, fig.cap= 'Scatter diagram of the energy demand (kW) versus energy usage (kWh)'}
# load data
ex51 = read.table("ex5-1.txt",header = T)
head(ex51)
# plot 
plot(ex51$x_.kWh., 
     ex51$y_.kW.,
     xlab = "Usage",
     ylab = "Demand")
```

As a starting point a simple linear regression model is assumed, and the least-squares fit is :

$$\hat y=-0.83130+0.00368x$$

```{r, fig.cap='Scatter diagram of the energy demand (kW) versus energy usage (kWh) with Simple Linear Model'}
# model
lm51 <- lm(ex51$y_.kW. ~ ex51$x_.kWh., data = ex51)
# summary (analysis of variance)
summary(lm51)
# plot 
plot(ex51$x_.kWh.,
     ex51$y_.kW.,
     xlab = "Usage",
     ylab = "Demand")
abline(lm51, col = "red")
```

For this model $R^2=0.7046$; that is about 70% of the variability in demand is accounted for by the straight-line fit to energy usage. The summary statistics do not reveal any obvious problems with this model. 

```{r, fig.cap= 'Plot of R-Student v.s fitted values'}
# R-student values versus fitter values fitted
plot(fitted(lm51),
     rstudent(lm51),
     ylab=TeX(r'($t_i$)'),
     xlab=TeX(r'($\hat{y}_i$)'),
     pch = 16);abline(0, 0,lty = 2)
```

Residuals form an outward-opening funnel, indicating that the error variance is increasing as energy consumption increases. A transformation may be helpful helpful in correcting this model inadequacy. To select the form of the transformation, note that the response variable y may be viewed as a "count" of the number of kilowatts used by a customer during a particular hour. The simplest probabilistic model for count data is the Poisson distribution. This suggests regressing $y^*=\sqrt{y}$ on x as a variance-stabilizing transformation. The resulting least-squares fit is : 

$$\hat y^*=0.5822+0.0009529x$$

```{r, fig.cap='Plot of R-student values versus fitted values for the transformed data'}
# y star 
ex51$ystar <- sqrt(ex51$y_.kW.)
# transformed model
lm51T <- lm(ex51$ystar ~ ex51$x_.kWh., data = ex51)# summary (analysis of variance)
summary(lm51T)
# R-student values versus fitted values for the transformed data
plot(fitted(lm51T),
     rstudent(lm51T),
     ylab=TeX(r'($t_i$)'),
     xlab=TeX(r'($\hat{y}^*_i$)'),
     pch = 16);abline(0, 0,lty = 2)
```

The impression from examining this plot is that the variance is stable; consequently, we conclude that the transformed model is adequate. 

Note that there is one suspiciously large residual (customer 26) and one customer whose energy usage is somewhat large (customer 50). The effect of these two points on the fit should be studied further before the model is released for use. 

# Example 5.2 : The Windmill Data 

A research engineer is investigating the use of a windmill to generate electricity. He has collected data on the DC Output from his windmill and the corresponding wind velocity. The data are plotted in the table and Figure 5 below. 

```{r, fig.cap='Plot of DC Output wind velocity x for the windmill data'}
# load data
ex52 <- read.table("ex5-2.txt",header = T)
head(ex52)
# plot
plot(ex52$WindVelocity_xi_mph,
     ex52$DCOutput_yi,
     xlab = "Wind Velocity, x",
     ylab = "DC Output, Y")
```

Inspection of the scatter diagram indicates that the relationship between DC output $(y)$ and wind velocity $(x)$ may be nonlinear. However, we initially fit a straight-line model to the data. The regression model is : 

$$\hat y=0.1309+0.2411x$$

```{r, fig.cap='Plot of DC Output wind velocity x for the windmill data with SLM'}
# model
lm52 <- lm(ex52$DCOutput_yi ~ ex52$WindVelocity_xi_mph, data = ex52)
# plot 
plot(ex52$WindVelocity_xi_mph,
     ex52$DCOutput_yi,
     xlab = "Wind Velocity, x",
     ylab = "DC Output, Y")
abline(lm52, col = "red")
# summary (analysis of variance)
summary(lm52)
# MSE <- c(crossprod(resid(lm52)))/length(resid(lm52));MSE
```

The summary statistics for this model are $R^2=0.8745$, and $F_0=160.26$ (the P-value is <0.0001). 

```{r}
# yhat and e 
ex52$fitted <- fitted(lm52)
ex52$resid <- resid(lm52)
ex52 %>% arrange(-desc(ex52$WindVelocity_xi_mph)) 
```

The observations are arranged in order of increasing wind speed. The residuals show a distinct pattern, that is, they move systematically from negative to positive and back to negative again as wind speed increases. 

```{r}
# plot 
plot(fitted(lm52),
     resid(lm52),
     ylab=TeX(r'($e_i$)'),
     xlab=TeX(r'($\hat{y}_i$)'),
     pch = 16);abline(0, 0,lty = 2)
```

This residual plot indicates model inadequacy and implies that the linear relationship has not captured all of the information in the wind speed variable. Note that the curvature was apparent in the earlier scatter diagram, is greatly amplified in the residual plot. 

Clearly some other model form must be considered. We might initially consider using a quadratic model such as :

$$y=\beta_0+\beta_1x+\beta_2x^2+\epsilon$$

to account for the curvature. However since the quadratic model will eventually bend downward as wind speed increases, it would not be appropriate for these data. A more reasonable model for windmill data that incorporates an upper asymptote would be :

$$y=\beta_0+\beta_1(\frac{1}{x})+\epsilon$$

```{r}
# transformation 
ex52$xstar <- 1/ex52$WindVelocity_xi_mph
# transformed model 
lm52T <- lm(ex52$DCOutput_yi ~ ex52$xstar, data = ex52)
# transformaion summary
summary(lm52T)
```

The fitted regression model is $$\hat y=2.9789-6.9345x'$$

The summary statistics for this model are$R^2=0.98$, and $F_0=1128$ (the p value is <0.0001).

```{r}
# R-student values versus fitted values for the transformed data
plot(fitted(lm52T),
     rstudent(lm52T),
     ylab=TeX(r'($t_i$)'),
     xlab=TeX(r'($\hat{y}_i$)'),
     pch = 16);abline(0, 0,lty = 2)
```

This plot does not reveal any serious problems. 

# Example 5.3 : The Electric Utility Data 
We use the Box-Cox procedure to select a variance-stabilizing transformation. The values of $SS_{Res}(\lambda)$ for various values are shown in the table ..


```{r}
boxcoxResult = boxcox(ex51$y_.kW. ~ ex51$x_.kWh., data = ex51, lambda = seq(-2,2,0.125))
```


