---
title: Using API's and tidycensus 
author: Randi Bolt
date: '2021-08-30'
slug: tidycensus-package
categories:
  - rstudio
tags:
  - data
  - rstudio
  - tutorial
  - api keys
  - population
  - census
---

* Register an API key at http://api.census.gov/data/key_signup.html 

* Create a .Renviron file in the main directory with "KEY=XXXXXXXXXXX". 

Note: this will not work with spaces on either side of the equal sign. 

Also note: tidycensus already has this utility worked into it (read ?census_api_key). They call their api key CENSUS_API_KEY (it is common for this key to be in all caps), so that is what I also called mine. This will be especially helpful in not mixing up API keys if I use other API keys in the future. 

```{r}
library(tidycensus)
# First time, reload your environment so you can use the key without restarting R.
# .../ tells the machine to go one folder outside the folder it is in
readRenviron("../../../.Renviron")
# You can check it with:
# Sys.getenv("CENSUS_API_KEY")
```

* load variables with load_variables(year, dataset, chache=T/F)

Read ?load_variables for various datasets and more information. 

Note that **label** shows the estimates by total, and then sex and age range. **concept** is by sex, then race, origins, and ancestry.  

```{r}
v15 <- load_variables(2019, "acs1")
v15
```

Let's only focus on the first line for now, "B01001_001" which should be the total estimates. Then we can use get_acs() to get data population data by state from the American Community Survey.

```{r}
get_acs(geography = "state", year = 2019, variable = "B01001_001")
```


We can get similar population estimates setting the variable = c("POP), with get_estimates. As well as "DENSITY"; for housing unit estimates, c("HUEST"); and for components of change estimates, c("BIRTHS", "DEATHS", "DOMESTICMIG", "INTERNATIONALMIG", "NATURALINC", "NETMIG", "RBIRTH", "RDEATH", "RDOMESTICMIG", "RINTERNATIONALMIG", "RNATURALINC", "RNETMIG").


```{r}
get_estimates(geography = "state", year = 2019, variable = c("POP"))
```

